---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# varnames

<!-- badges: start -->
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/milanwiedemann/varnames?branch=master&svg=true)](https://ci.appveyor.com/project/milanwiedemann/varnames)
[![Travis build status](https://travis-ci.com/milanwiedemann/varnames.svg?branch=master)](https://travis-ci.com/milanwiedemann/varnames)
<!-- badges: end -->

The goal of `varnames` is to help with creating consistent variable names.
Good variable names have many advantages, particularly when managing larger datasets.

Consistent names make also make it easier to transform data sets to "tidy data", see [Wickham (2013)](https://www.jstatsoft.org/article/view/v059i10).
I'll show some examples how consistent variable names can be really powerful because it's easier to automate things and use [Regular Expressions](https://stringr.tidyverse.org/articles/regular-expressions.html).


## Installation

This is work in progress, the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("milanwiedemann/varnames")
```
## Example

### Base R

For most cases there might be a very easy way to create consistent variable names simply by using the `paste()` function. 
This function comes with base R which means that no further packages are needed and it is very flexible!

```{r example-paste}
# Create vector of a measure with 3 time points (t)
paste("measure_", "t", 1:3, sep = "")
```

Sometimes more than one specification need to be added to a variable name.
This is still possible using base function but it gets a bit tricky so I build functions that help with this.
 
### varnames R package 

Similar to the example above, the `create_var_names()` function creates a vector (or list) of a measure with 3 time points ("t") specified in the arguments `str = "t"` and `n = 3`.

```{r example-create}
library(varnames)
library(tidyverse)

# Create vector of a measure with 3 time points (t)
create_var_names(var_name = "measure", str = "t", n = 3, unlist = T)
```

The `add_specifier()` function adds a further specification to the variables created in the step before, here `"i"` for item.
Depending on what output format is most suitable for further use, this function can return the output in different ways. 

The example below adds two items "i" for each time point "t" and returns a list that is sorted by the output format from the previous function using `sort = "previous"`. 
This creates a list with 3 vectors, one for each time point specified in the previous function.

```{r example-add-1}
create_var_names(var_name = "measure", str = "t", n = 3) %>% 
  add_specifier(str = "i", n = 2, sort = "previous")
```

The output list can also be sorted by the new specifyer that was added by using the argument `sort = "current"`.
Now a list with two vectors is returned, one vector for each indicator that was added.

```{r example-add-2}
create_var_names(var_name = "measure", str = "t", n = 3) %>% 
  add_specifier(str = "i", n = 2, sort = "current")
```
A list can be returned using the argument `unlist = TRUE`. 

```{r example-add-3}
create_var_names(var_name = "measure", str = "t", n = 3) %>% 
  add_specifier(str = "i", n = 2, sort = "current", unlist = TRUE)
```

The order of the vector can still be changed when using `unlist = TRUE` by changing the sort argument.

```{r example-add-4}
create_var_names(var_name = "measure", str = "t", n = 3) %>% 
  add_specifier(str = "i", n = 2, sort = "previous", unlist = TRUE)
```

## TODOs

- Say something about "tidy" data [@Wickham2013](https://www.jstatsoft.org/article/view/v059i10)
- In the examples below I'm showing how data stored in these variable names can be transformend into "long" data.
- Add Why is this so useful? Show some examples, regular expressions automation etc
- Add rename functions
- Currently this only works when adding one specifyer, I think more would be cool, I might need help with this 
